%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% BENELEARN 2013 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following line _only_ if you're still using LaTeX 2.09.
%\documentstyle[benelearn2013,epsf,mlapa]{article}
% If you rely on Latex2e packages, like most moden people use this:
\documentclass{article}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''

\usepackage{benelearn2013}

% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Appearing in''
%\usepackage[accepted]{benelearn2013}

% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

% For citations
\usepackage{mlapa}

% The \benelearntitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\benelearntitlerunning{Formatting Instructions for BENELEARN 2013 Abstracts}

\begin{document}

\twocolumn[
\benelearntitle{Formatting Instructions for BENELEARN 2013 Abstracts \\
           \texttt{http://www.benelearn2013.org/}}

 \benelearnauthor{Your Name}{email@yourdomain.edu}
 \benelearnaddress{Your Fantastic Institute,
             van Goghstraat 7, 2387, Baarle-Hertog, Belgium}
 \benelearnauthor{Your CoAuthor's Name}{email@coauthordomain.edu}
 \benelearnaddress{Their Fantastic Institute,
             Magrittelaan 42, 5110 AB Baarle-Nassau, The Netherlands}
\benelearnaddress{{\bf Keywords}: submission information, deadline data,
	formatting information
          }
\vskip 0.3in
]

Abstracts and papers should be submitted electronically, no later than March 12, 2013. Notification of acceptance is on April 16.

BENELEARN 2013 will rely exclusively on electronic formats for submission and review. Authors are expected to write their paper with the \textbf{\LaTeX} editor, for which electronic templates are accessible at the BENELEARN website. If you are using the templates, the formatting instructions below will already be enforced.

References must include page numbers whenever possible and be as complete as possible. Place multiple citations in chronological order. For abstracts dealing with published work, please indicate this in the text and provide the appropriate references. The review process will be single blind, so also mention your name and affiliation. Please consult the template for full papers to format references. If the authors' names are included in the sentence, place only
the year in parentheses, for example when referencing Rob Schapire's seminal result \yrcite{schapire90}. Otherwise place the entire reference in parentheses with the authors and year separated by a comma \cite{schapire90,kearns89}.

An abstract should not be longer than one page. Submission to BENELEARN 2013 will be entirely electronic, through Easychair.  The URL and information about the submission process will appear on the conference web site. To ensure our ability to print submissions, authors must provide their manuscripts in PDF format.

Authors need to pay close attention to the typefaces used. More information about this can be found in the template for full papers. Problems can be avoided by using the \textbf{pdflatex} program instead of straight \LaTeX. This program avoids the Type-3 font problem, however you must ensure that all of the fonts are embedded (use {\tt
pdffonts}).

\section*{Acknowledgments}
\begin{scriptsize}
{\footnotesize •}
\end{scriptsize}We acknowledge that this template is a modification of the template for BeneLearn 2012.

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite
%\nocite{zinkevich03}

\bibliography{example_paper}
\bibliographystyle{mlapa}

\end{document}

% Version History
% Document updated by Florian Kunneman for BENELEARN 2013
% Document updated by Willem Waegeman for BeneLearn 2012
% Document updated by Peter van der Putten for BeneLearn 2011
% This document was modified by Kurt Driessens for BeneLearn 2010.
% Before it was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This 2009 version was
% created by Kiri Wagstaff, slightly modified from Sam Roweis's 2008
% version, which is slightly modified from Prasad Tadepalli's 2007
% version which is a lightly changed version of the previous year's
% version by Andrew Moore, which was in turn edited from those of
% Kristian Kersting and Codrina Lauth. Alex Smola contributed to the
% algorithmic style files.


